Reconstructing the neutron-star equation of state with gravitational-wave detectors from a realistic population of inspiralling binary neutron stars

### 2014-08-31 paper.tex
Finished paper comments

As an alternative to the PN approximation, the effective one body (EOB) formalism, which uses various techniques to re-sum the PN series, may converge more rapidly to the true binary waveform. EOB waveforms have been shown to accurately reproduce numerical binary black hole (BBH) waveforms. For example, Damour {\it et al.}~\cite{DamourNagarBernuzzi2013} compared a recent EOB implementation with nonspinning BBH simulations of the last $\sim 30$ GW cycles and found a phase difference of $<0.1$~radians after fitting the unknown 5PN contribution in the EOB radial potential to the numerical data. Since the tidal contribution to the waveform over this same interval is usually more than a radian, the current EOB waveform may be accurate enough for removing systematic errors due to uncertainties in the point-particle model. Tidal interactions have also been calculated in the EOB formalism for the first few multipoles to 2PN order in the EOB radial potential~\cite{BiniDamourFaye2012}, and comparisons with numerical BNS simulations have shown that EOB waveforms, after calibration, are consistent with the numerical waveforms~\cite{BernuzziNagarThierfelder2012, HotokazakaKyutokuShibata2013} \les{(Well of course they do, they have been calibrated to match lol!  The problem with these waveforms is in the ``in between'' range.  Basically, EOB are like PN waveforms in the early inspiral and calibrated to NR in the very late inspiral.  But, since current NR waveforms are only evolved out for the last several cycles, there might still be significant phase uncertainty before these last several cycles, in the range where the PN approximations break down.  This is my understanding at least.  However, I do still think that these are the best available waveforms.  If only they were in LAL!)}{}. Unfortunately, due to difficulties from hydrodynamics, BNS \les{NR}{} codes do not yet have the same level of accuracy as BBH codes.

@@

The original intent of the 4-parameter piecewise polytrope was to provide a reasonable fit to a wide range of EOS models with a small number of parameters that have an intuitive meaning~\cite{ReadLackey2009}. However, it appears that for directly measuring the EOS with inspiral observations, a more sophisticated EOS will be needed. Two possibilities mentioned in Section~\ref{sec:eosparam}, include the parameterization in Ref.~\cite{SteinerLattimerBrown2010} which uses variable density intervals as well as the spectral fit of Ref.~\cite{Lindblom2010}. However, the optimal parameterization for recovering the EOS and useful nuclear parameters from GW data remains an open question.

@@

BL sincerely thanks Walter Del Pozzo and Tyson Littenberg for advice, many helpful discussions, and correcting misunderstandings in the implementation of Bayesian techniques, as well as Will Farr, Tjonnie Li, Chris Van Den Broeck, and Andrew Steiner. BL also had several informative discussions at the University of Washington's Institute for Nuclear Theory Program 14-2a. The authors would also like to thank Evan Ochsner and Jolien Creighton for advice and for generating the parameters of the BNS population used here. This work was supported by NSF Grants PHY-1305682, and PHY-\red{xxxxxxx [[Les, fill this in]]}. Simulations were run on the Nemo cluster at the University of Wisconsin--Milwaukee and the Orbital cluster at Princeton University.

>>>>

As an alternative to the PN approximation, the effective one body (EOB) formalism, which uses various techniques to re-sum the PN series, may converge more rapidly to the true binary waveform. EOB waveforms have been shown to accurately reproduce numerical binary black hole (BBH) waveforms. For example, Damour {\it et al.}~\cite{DamourNagarBernuzzi2013} compared a recent EOB implementation with nonspinning BBH simulations of the last $\sim 30$ GW cycles and found a phase difference of $<0.1$~radians after fitting the unknown 5PN contribution in the EOB radial potential to the numerical data. Since the tidal contribution to the waveform over this same interval is usually more than a radian, the current EOB waveform may be accurate enough for removing systematic errors due to uncertainties in the point-particle model. Tidal interactions have also been calculated in the EOB formalism for the first few multipoles to 2PN order in the EOB radial potential~\cite{BiniDamourFaye2012}, and comparisons with numerical BNS simulations have shown that EOB waveforms, after calibration, are consistent with the numerical waveforms~\cite{BernuzziNagarThierfelder2012, HotokazakaKyutokuShibata2013} \les{(Well of course they do, they have been calibrated to match lol!  The problem with these waveforms is in the ``in between'' range.  Basically, EOB are like PN waveforms in the early inspiral and calibrated to NR in the very late inspiral.  But, since current NR waveforms are only evolved out for the last several cycles, which is far later than we can trust PN waveforms, there might still be significant de-phasing before these last several cycles, which will still bias the recovery of $\tilde\Lambda$.  This is my understanding at least, though I will admit that I'm no expert on EOB waveforms.  However, I do still think that these are the best available waveforms.  If only they were in LAL!)}{}. Unfortunately, due to difficulties from hydrodynamics, BNS \les{NR}{} codes do not yet have the same level of accuracy as BBH codes.

@@

The original intent of the 4-parameter piecewise polytrope was to provide a reasonable fit to a wide range of EOS models with a small number of parameters that have an intuitive meaning~\cite{ReadLackey2009}. However, it appears that for directly measuring the EOS with inspiral observations, a more sophisticated EOS will be needed. Two possibilities mentioned in Section~\ref{sec:eosparam}, include the parameterization in Ref.~\cite{SteinerLattimerBrown2010} which uses variable density intervals \les{or }{as well as} \les{(the ``as well as'' made me think that the parameterization in Ref.~\cite{SteinerLattimerBrown2010} used variable density intervals and a spectral fit.)}{} the spectral fit of Ref.~\cite{Lindblom2010}. However, the optimal parameterization for recovering the EOS and useful nuclear parameters from GW data remains an open question.

@@

BL sincerely thanks Walter Del Pozzo and Tyson Littenberg for advice, many helpful discussions, and correcting misunderstandings in the implementation of Bayesian techniques, as well as Will Farr, Tjonnie Li, Chris Van Den Broeck, and Andrew Steiner. BL also had several informative discussions at the University of Washington's Institute for Nuclear Theory Program 14-2a. The authors would also like to thank Evan Ochsner and Jolien Creighton for advice and for generating the parameters of the BNS population used here. This work was supported by NSF Grants PHY-1305682, \les{PHY-0970074, and PHY-0955929}{}. Simulations were run on the Nemo cluster, \les{which is supported by the NSF Grant PHY-0923409}{}, at the University of Wisconsin--Milwaukee and the Orbital cluster at Princeton University.
